Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-08-06 surfaces-write-up/surface-write-up.tex
some todos like F_mu as iota circ psi

\todo{Incorporate the following section into the $\bp^{m}$ section
\\
Peter: I'm not sure what this refers to...  Can you be more specific}

@@

\todo{Aaron: weren't we going to take this section out and move it into the later sections?}

@@

We now present some preliminary results which will be used to prove Theorems \ref{thm:proj-generators-relations} and \ref{thm:hirz-generators-relations} in Sections \ref{sec:proj} and \ref{sec:hirz} respectively.
%Now we give a result describing the generators of a cone via its
%extremal rays.
%\todo{write more transition}

@@

\noindent
has a minimal presentation over $R_{D'}$ consisting of the $\sum_{i = 0}^{r}
{{m + c_i - 1} \choose {c_i}}$ generators $F_i^{\vec{v}} := \frac{u^{d_i}
x^{\vec{v}}}{x_k^{c_i}}$ where $1 \leq i \leq r$, $\vec{v} \in \bz_{\geq 0}^{m + 1}$
with $\deg \vec v = c_i$. Furthermore, it has
relations generated by the following two forms:
\begin{enumerate}
	\item For each $(i, j)$ with $j \geq i + 2$ and each $\vec{v} \in \mss_i,
\vec{w} \in \mss_j$, there is a relation of the form: either
\begin{flalign*}
	&G_{i, j}^{\vec{v}, \vec{w}} = F_i^{\vec{v}} F_j^{\vec{w}}
	- \prod_{\vec{y} \in \mss_{h_{i, j}}} (F_{h_{i, j}}^{\vec{y}})
	^{g_{\vec{y}}} &(i < h_{i, j} < j) \\
	\text{or} \\
	&G_{i, j}^{\vec{v}, \vec{w}} = F_i^{\vec{v}} F_j^{\vec{w}}
	- \prod_{\vec{y} \in \mss_{h_{i, j}}} (F_{h_{i, j}}^{\vec{y}})^{g_{\vec{y}}}
	\cdot \prod_{ \vec{z} \in
	\mss_{h_{i, j} + 1}} (F_{k_{i, j} + 1}^{\vec{z}})^{g'_{\vec{z}}}
	&(i < h_{i, j} < h_{i, j} + 1 < j).
\end{flalign*}
\todo{Peter: Can you clarify when these two cases occur: i.e. does the first 
only occur if $j = h_{i,j}+1$; otherwise, in cases when $i< h_{i,j} < h_{i,j} + 1< 
j$, do we get relations of both forms? \\
Robin: It purely depends on whether the product $F_i^{\vec{v}} F_j^{\vec{w}} 
$ happens to lie on the ray given by $\frac{c_{h_{i, j}}}{d_{h_{i, j}}}$ which is 
not clearly determined by $i$ or $j$ from what I can see... Unless you see 
something else. Only one relation exists for each $(i, j)$ and it is exactly one 
of the two forms. I figured that the ``either'' made that clear, no?}
	\item For each $(i, j)$ with
$j = i$ or $j = i + 1$ and each $\vec{v} \in \mss_i, \vec{w} \in
\mss_j$ with $\vec{v} \not\prec \vec{w}$ (see Definition
~\ref{defn:vec-order}) there is a relation of the form
\begin{align*}
	&L_{i, j}^{\vec{v}, \vec{w}} = F_i^{\vec{v}} F_j^{\vec{w}}
	- F_i^{\vec{y}} F_j^{\vec{z}} \\
\end{align*}

@@

If you want, can you tell me exactly which variables would be helpful in understanding the proof to know their dependencies?
}
 whenever $j > 4$.  To bound the degree of generation of these relations, we bound the degree of generation of the ideal $(z_i - \beta_i) \cap \ker(\psi)$ for each $i$.

@@

Let $M_i$ be the set of monomials terms of $\beta_i = \beta_i(f_1, \ldots f_4)$.  Let $\mu = \prod_{j=1}^4
{f_j}^{h_j}\in M_i$ and consider the lattice point 
\[
	\sigma_\mu = (d, c_1 + h_1, \ldots, c_4+h_4, c_5, \ldots, c_n).
\]
where $d = \deg \sigma$.
Define 
\[
	E_\mu = \sigma_\mu - \lambda.
\]
A quick computation deduces that $E_\mu$ lies in $\Sigma$ 
%(using the condition on the $r_{i,j}$'s and $r_{j,i}s$) \todo{should one of these be $r_i$'s?}.  \\ Peter: No, but I'm not sure if the parenthetical is helpful anyway.  $r_i$ isn't define or used (hopefully...) anywhere.) 
and has the same degree as $E$.

Let $F = \Psi(E) \in R_D$
\todo{Aaron: Didn't we agree you would write these as $\psi(E)$ or something like that? 
\\
Peter: Okay, I'm still a little unhappy with with the identification of lattice points in $\Sigma$ with monomials (since $\Psi$ maps monomials in a polynomial ring to monomials in $R_D$), but I made the change} and $F_\mu = \psi(E_\mu) \in R_D$ for each $\mu \in M_i$.  
Then $F - \sum_{\mu \in M_i} F_\mu = 0$ is a relation %in degree
%\[
%	\deg(F) = \deg(E)
%\]
which divides $(z_i - \beta_i) \prod_{j = 1}^n z_j^{c_j}$. 
We can bound $\deg(F)$ (which equals $\deg(F_\mu)$ for all $\mu$) by
\[
	\deg(F) = \deg(E) \le \rho + \begin{cases}
	\ell_i \gcd(\bida_i, \bidb_i)	&\mbox{ if } i \in \Te \\
	\max_{j \in \Tm} \bigl(\ell_{i, j} (\bida_i \bidb_j - \bida_j \bidb_i)\bigr)
	&\mbox{ if } i \in \Tp \\
	\max_{j \in \Tp} \bigl(\ell_{j, i} (\bida_j \bidb_ i - \bida_i \bidb_j) \bigr)
	&\mbox{ if } i \in \Tm. \end{cases}
	\]

>>>>

\todo{Aaron: Incorporate the following section into the $\bp^{m}$ section
\\
Peter: I'm not sure what this refers to...  Can you be more specific \\

Aaron: A couple of weeks ago, we agreed that we would remove the general lemmas section, and just
merge it with the $\bp^m$ section. So, I was saying that we should do this.}

@@

@@

We now present some preliminary results which will be used to prove Theorems \ref{thm:proj-generators-relations} and \ref{thm:hirz-generators-relations} in Sections \ref{sec:proj} and \ref{sec:hirz} respectively.

@@

\noindent
has a minimal presentation over $R_{D'}$ consisting of the $\sum_{i = 0}^{r}
{{m + c_i - 1} \choose {c_i}}$ generators $F_i^{\vec{v}} := \frac{u^{d_i}
x^{\vec{v}}}{x_k^{c_i}}$ where $1 \leq i \leq r$, $\vec{v} \in \bz_{\geq 0}^{m + 1}$
with $\deg \vec v = c_i$. Furthermore, it has
relations generated by the following two forms:
\begin{enumerate}
	\item For each $(i, j)$ with $j \geq i + 2$ and each $\vec{v} \in \mss_i,
\vec{w} \in \mss_j$, there is a relation of the form either
\begin{flalign*}
	&G_{i, j}^{\vec{v}, \vec{w}} = F_i^{\vec{v}} F_j^{\vec{w}}
	- \prod_{\vec{y} \in \mss_{h_{i, j}}} (F_{h_{i, j}}^{\vec{y}})
	^{g_{\vec{y}}} &(i < h_{i, j} < j) \\
	\text{or} \\
	&G_{i, j}^{\vec{v}, \vec{w}} = F_i^{\vec{v}} F_j^{\vec{w}}
	- \prod_{\vec{y} \in \mss_{h_{i, j}}} (F_{h_{i, j}}^{\vec{y}})^{g_{\vec{y}}}
	\cdot \prod_{ \vec{z} \in
	\mss_{h_{i, j} + 1}} (F_{k_{i, j} + 1}^{\vec{z}})^{g'_{\vec{z}}}
	&(i < h_{i, j} < h_{i, j} + 1 < j).
\end{flalign*}
\todo{Peter: Can you clarify when these two cases occur: i.e. does the first 
only occur if $j = h_{i,j}+1$; otherwise, in cases when $i< h_{i,j} < h_{i,j} + 1< 
j$, do we get relations of both forms? \\
Robin: It purely depends on whether the product $F_i^{\vec{v}} F_j^{\vec{w}} 
$ happens to lie on the ray given by $\frac{c_{h_{i, j}}}{d_{h_{i, j}}}$ which is 
not clearly determined by $i$ or $j$ from what I can see... Unless you see 
something else. Only one relation exists for each $(i, j)$ and it is exactly one 
of the two forms. I figured that the ``either'' made that clear, no?}
	\item For each $(i, j)$ with
$j = i$ or $j = i + 1$ and each $\vec{v} \in \mss_i, \vec{w} \in
\mss_j$ with $\vec{v} \not\prec \vec{w}$ (see Definition
~\ref{defn:vec-order}) there is a relation of the form
\begin{align*}
	&L_{i, j}^{\vec{v}, \vec{w}} = F_i^{\vec{v}} F_j^{\vec{w}}
	- F_i^{\vec{y}} F_j^{\vec{z}} \\
\end{align*}

@@

If you want, can you tell me exactly which variables would be helpful in
understanding the proof to know their dependencies?

Aaron: Ok, I'll stop bugging you about it. I just thought we
should add it because I didn't understand that it depended
on $i$ when I read it. Maybe we can see what David
has to say about it when he reads it?
Also, I don't see why calling $s_{j,k}$ by $s_{j,k,\sigma}$ is cumbersome,
that seems fine to me.
}

 whenever $j > 4$.  To bound the degree of generation of these relations, we bound the degree of generation of the ideal $(z_i - \beta_i) \cap \ker(\psi)$ for each $i$.

@@

Let $M_i$ be the set of monomials terms of $\beta_i = \beta_i(f_1, \ldots f_4)$.  Let $\mu = \prod_{j=1}^4
{f_j}^{h_j}\in M_i$ and consider the lattice point 
\[
	\sigma_\mu = (d, c_1 + h_1, \ldots, c_4+h_4, c_5, \ldots, c_n).
\]
where $d = \deg \sigma$.
Define 
\[
	E_\mu = \sigma_\mu - \lambda.
\]
A quick computation deduces that $E_\mu$ lies in $\Sigma$ 
and has the same degree as $E$.

Let $F = \Psi(E) \in R_D$
\todo{Aaron: Didn't we agree you would write these as $\psi(E)$ or something like that? 
\\
Peter: Okay, I'm still a little unhappy with with the identification of lattice
points in $\Sigma$ with monomials (since $\Psi$ maps monomials in a polynomial
ring to monomials in $R_D$), but I made the change
Aaron: Thanks. If you're unhappy, you can always just define the map $\iota:\Sigma \rightarrow \bk[\Sigma]$ and say $F_\mu = \iota \circ \psi (E_\mu)$}
 and $F_\mu = \psi(E_\mu)
\in R_D$ for each $\mu \in M_i$.  
Then $F - \sum_{\mu \in M_i} F_\mu = 0$ is a relation %in degree
%\[
%	\deg(F) = \deg(E)
%\]
which divides $(z_i - \beta_i) \prod_{j = 1}^n z_j^{c_j}$. 
We can bound $\deg(F)$ (which equals $\deg(F_\mu)$ for all $\mu$) by
\[
	\deg(F) = \deg(E) \le \rho + \begin{cases}
	\ell_i \gcd(\bida_i, \bidb_i)	&\mbox{ if } i \in \Te \\
	\max_{j \in \Tm} \bigl(\ell_{i, j} (\bida_i \bidb_j - \bida_j \bidb_i)\bigr)
	&\mbox{ if } i \in \Tp \\
	\max_{j \in \Tp} \bigl(\ell_{j, i} (\bida_j \bidb_ i - \bida_i \bidb_j) \bigr)
	&\mbox{ if } i \in \Tm. \end{cases}
	\]
